---
title: "COI EtOH neutralization"
output: html_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message=FALSE, warning=FALSE)
knitr::is_latex_output()
#output: html_document:
#    keep_md: true
library(phyloseq)
library(vegan)
library(ggplot2)
library(dplyr)
library(tidyr)
library(clustsig)
library(ggdendro)
library(gridExtra)
library(stringr)
library(dendextend)
library(Biostrings)
library(insect)
library(ape)
library(seqinr)
library(decontam)
library(lmodel2)
library(mgcv)
library(goeveg)
library(metagMisc)
library(dendextend)
library(magrittr)
library(rstatix)
library(lemon)
theme_update(plot.title = element_text(hjust = 0.5))
```

```{r import COI data}
in_biom <- import_biom("~/Documents/Chapter4/etoh_nh4oh/exported-feature-table/feature-table_json.biom")
#sample_names(in_biom)

md <- import_qiime_sample_data("~/Documents/Chapter4/etoh_nh4oh/metadata_2.txt") #### 
in_biom <- merge_phyloseq(in_biom, md)

#assigned through midori web reference http://reference-midori.info/server.php
#with RDP classifier, COI, unique, confidence cutoff 0.8
taxcsv <- read.csv("~/Documents/Chapter4/etoh_nh4oh/RDPClassifier_out/20201015225551_24.43.252.202_usga_classified.txt", sep="\t", header = F)

###this separates the taxonomy from the confidence 
#taxonomy <- taxcsv[,c(6, 9, 12, 15, 18, 21, 24)]
#taxonomy_conf <- taxcsv[,c(8, 11, 14, 17, 20, 23, 26)]
#rownames(taxonomy) <- taxcsv$V1
#rownames(taxonomy_conf) <- taxcsv$V1
#taxfull <- merge(taxonomy, taxonomy_conf)

###this filters the taxonomy from the confidence
mylist <- list() #create an empty list
k <- 1
for (i in c(6, 9, 12, 15, 18, 21, 24)){
  j <- i + 2
  confidence <- taxcsv[,j]
  #truefalse <- truefalse > 0.8
  assignment <- taxcsv[,i]
  assignment[confidence < 0.8] <- NA
  mylist[[k]] <- assignment
  k <- k+1
}
taxonomy <- do.call("cbind",mylist) #combine all vectors into a matrix
rownames(taxonomy) <- taxcsv$V1

#convert to phyloseq objects
in_biom_tax <- tax_table(taxonomy)
taxa_names(in_biom_tax) <- rownames(taxonomy)

#### combine filtered phyloseq object and taxonomy infor
in_biom <- merge_phyloseq(in_biom, in_biom_tax)

COI <- in_biom

save(COI, file = "~/Documents/Chapter4/etoh_nh4oh/COI_v1biom_etohnh4oh_midoritax.Rdat")

rm(list=c("in_biom", "md", "taxonomy", "taxcsv", "mylist", "taxonomy"))
```

```{r}
## ok - need to  combine the non-quant subsamples relative to their original biomass
sample_data(COI)
scratch <- rarefy_even_depth(COI)
size_factors <- sample_data(scratch)$biomass #get original biomass
size_factors2 <- size_factors/c(6.957, 6.957, 11.041, 3.781, 5.475, 9.993, 9.993, 4.452, 6.957, 11.041, 3.781, 5.475, 9.993, 9.993, 4.452, 6.957) #divide by total biomass for each sample
mat <- sweep(otu_table(scratch), 1 + taxa_are_rows(scratch), size_factors2, FUN = `/`) #adjust sequencing depth to reflect fraction of original biomass
mat <- round(mat)
scratch0 <- scratch #add adjusted otu table back into phyloseq object
otu_table(scratch0) <- otu_table(mat, taxa_are_rows = taxa_are_rows(scratch)) 
otu_table(scratch0)
sample_data(scratch0)$mergebiomass <- paste(sample_data(scratch0)$sampleunique, sample_data(scratch0)$rep)
COI_mergedbiomass <- merge_samples(scratch0, group = "mergebiomass") #merge 

#get sample data out of sample names (coerced to NAs by merge)
sample_data(COI_mergedbiomass)$sampleid <- strsplit(sample_names(COI_mergedbiomass), " ", ) %>% sapply(extract2, 1)

COI_proportional_merge <- merge_samples(COI_mergedbiomass, group = "sampleid") #merge replicates
sample_data(COI_proportional_merge)$TowLocation <- strsplit(sample_names(COI_proportional_merge), "-", ) %>% sapply(extract2, 1)
sample_data(COI_proportional_merge)$TowTime <- strsplit(sample_names(COI_proportional_merge), "-", ) %>% sapply(extract2, 2)
sample_data(COI_proportional_merge)$Preservative <- strsplit(sample_names(COI_proportional_merge), "-", ) %>% sapply(extract2, 3)


```

```{r combine paired samples}
#includes both non-quant splits from extraction process and duplicate sequencing 
COI_merged <- merge_samples(COI, group = "sampleunique")
sample_data(COI_merged)$preservative <- c("Buffered", "Unbuffered", "Buffered", "Unbuffered", "Buffered", "Unbuffered")
sample_data(COI_merged)$sampleid <- sample_names(COI_merged)
sample_data(COI_merged)$towlocation <- c("CCE2-D", "CCE2-D", "CCE2-N", "CCE2-N", "LJC", "LJC")
sample_data(COI_merged)

```


```{r buff v unbuff richness}
#plot_bar(COI_merged, fill = "ta3") check what groups we have
COI_merged_rarefied <- (COI_proportional_merge)
sample_data(COI_merged_rarefied)$total_richness <- estimate_richness(COI_merged_rarefied, measures = "Observed")$Observed
total <- COI_merged_rarefied
malacostraca <- subset_taxa(COI_merged_rarefied, ta3 == "Malacostraca")
maxillopoda <- subset_taxa(COI_merged_rarefied, ta3 == "Maxillopoda")
ostracoda <- subset_taxa(COI_merged_rarefied, ta3 == "Ostracoda")
polychaeta <- subset_taxa(COI_merged_rarefied, ta3 == "Polychaeta")
hydrozoa <- subset_taxa(COI_merged_rarefied, ta3 == "Hydrozoa")
ophiuroidea <- subset_taxa(COI_merged_rarefied, ta3 == "Ophiuroidea")
gymnolaemata <- subset_taxa(COI_merged_rarefied, ta3 == "Gymnolaemata")
gastropoda <- subset_taxa(COI_merged_rarefied, ta3 == "Gastropoda")
brachiopoda <- subset_taxa(COI_merged_rarefied, ta3 == "Branchiopoda")
actinopteri <- subset_taxa(COI_merged_rarefied, ta3 == "Actinopteri")
insecta <- subset_taxa(COI_merged_rarefied, ta3 == "Insecta")
cephalopoda <- subset_taxa(COI_merged_rarefied, ta3 == "Cephalopoda")
polychaeta <- subset_taxa(COI_merged_rarefied, ta3 == "Polychaeta")
echinoidea <- subset_taxa(COI_merged_rarefied, ta3 == "Echinoidea")


sample_data(COI_merged_rarefied)$Malacostracans <- estimate_richness(malacostraca, measures = "Observed")$Observed
sample_data(COI_merged_rarefied)$Maxillopods <- estimate_richness(maxillopoda, measures = "Observed")$Observed
sample_data(COI_merged_rarefied)$Ostracods <- estimate_richness(ostracoda, measures = "Observed")$Observed
sample_data(COI_merged_rarefied)$Polychaetes <- estimate_richness(polychaeta, measures = "Observed")$Observed
sample_data(COI_merged_rarefied)$Hydrozoans <- estimate_richness(hydrozoa, measures = "Observed")$Observed
sample_data(COI_merged_rarefied)$Ophiuroids <- estimate_richness(ophiuroidea, measures = "Observed")$Observed
sample_data(COI_merged_rarefied)$Bryozoans <- estimate_richness(gymnolaemata, measures = "Observed")$Observed
sample_data(COI_merged_rarefied)$Gastropods <- estimate_richness(gastropoda, measures = "Observed")$Observed
sample_data(COI_merged_rarefied)$Brachiopods <- estimate_richness(brachiopoda, measures = "Observed")$Observed
sample_data(COI_merged_rarefied)$Fishes <- estimate_richness(actinopteri, measures = "Observed")$Observed
#sample_data(COI_merged_rarefied)$Insects <- estimate_richness(insecta, measures = "Observed")$Observed
sample_data(COI_merged_rarefied)$Cephalopods <- estimate_richness(cephalopoda, measures = "Observed")$Observed
sample_data(COI_merged_rarefied)$Polychaetes <- estimate_richness(polychaeta, measures = "Observed")$Observed
sample_data(COI_merged_rarefied)$Echinoderms <- estimate_richness(echinoidea, measures = "Observed")$Observed

df_new <- data.frame(sample_data(COI_merged_rarefied))
df_new <- df_new[,2:ncol(df_new)]

df_pivoted1_new <- df_new %>% pivot_longer(cols = colnames(df_new)[c(15:27)],
                    names_to = "group",
                    values_to = "observed") %>%
  ungroup() 

df_pivoted2_new <- df_pivoted1_new %>%
  pivot_wider(id_cols = c("TowLocation", "TowTime", "group"),
    names_from = c("Preservative"),
              values_from = "observed") %>%
  ungroup()
df_pivoted2_new$Sample <- paste(df_pivoted2_new$TowLocation, df_pivoted2_new$TowTime)
df_pivoted2_new$Sample[df_pivoted2_new$Sample == "LJC Morning"] <- "LJC Day"



pairwise.wilcox.test(df_pivoted1_new$observed, df_pivoted1_new$Preservative, paired = T)

p <- ggplot(data = df_pivoted2_new, aes(x = unbuff, y = buffered, color = group, shape = Sample)) + 
  geom_point(size=.8) + 
  guides(size = FALSE, color=guide_legend(keywidth=2.5,
                 keyheight=2.5,
                 default.unit="mm"), 
          shape=guide_legend(keywidth=2.5,
                 keyheight=2.5,
                 default.unit="mm"),
         legend.key = element_rect(size = 2.5, fill = "white", colour = NA)) +
  geom_abline(intercept=0, slope=1, size = 0.1) +
  theme_bw(base_size = 6) + 
  xlab("Richness in Untreated Samples") + 
  ylab("Richness in Neutralized Samples ") + 
  coord_fixed(ratio = 1, xlim = NULL, ylim = NULL, expand = TRUE, clip = "on") +
  labs(size = NULL, shape = "Sample", color = "Taxon") +
  scale_color_discrete(name = "Taxon", labels = c("Brachiopods", "Bryozoans", "Cephalopods", "Echinoderms", "Fishes", "Gastropods", "Hydrozoans", "Malacostracans", "Maxillopods", "Ophiuroids", "Ostracods", "Polychaetes", "Total Richness"))
#  ylim(0,65) + 
#  xlim(0,65)
p

p1 <- p 
p2 <- p + ylim(c(0, 17)) + xlim(c(0,17)) + labs(size = NULL, shape = NULL, color = NULL) + theme(legend.position = "none") + ylab(NULL) + xlab(NULL) + geom_point(size = .5) 


tiff('~/Documents/Chapter4/Oct2020_ICES/plots/Figure4_v2_2020-11-17.tiff', width=85,height=60, units = "mm", res = 300)
p1 + annotation_custom(ggplotGrob(p2), xmin = 115, xmax = 240, 
                       ymin = -8.5, ymax = 110) +
 theme_bw(base_size = 6) +
  theme(axis.text.x = element_text(color = "black", size = 6),
        axis.text.y = element_text(color = "black", size = 6),  
        axis.title.x = element_text(color = "black", size = 7),
        axis.title.y = element_text(color = "black", size = 7),
        legend.spacing.y = unit(1.0, 'mm'))
dev.off()



```


```{r simprof}
scratch <- COI_proportional_merge
scratch <-  transform_sample_counts(scratch, function(x) x/sum(x))
sample_names(scratch) <- c("CCE2-D Neutralized", "CCE2-D Untreated", "CCE2-N Neutralized", "CCE2-N Untreated", "LJC Neutralized", "LJC Untreated")
res2 <- simprof(data=(data.frame(otu_table(scratch))), 
method.distance="braycurtis", num.expected = 1000, num.simulated = 999, silent = FALSE, increment = 100)
# Graph the result
#pl.color <- simprof.plot(res2)
```

```{r pretty plot simprof}

#tiff('plots/Figure5_v3-2020-11-13.tiff', width=85,height=90, units = "mm", res = 300) 
#par(mar=c(10,3,1,1))  # Increase bottom margin to have the complete label
#simprof.plot(res2, leafcolors = c("red", "blue"), siglinetype = 2, lwd = 0.8)
#dev.off()


```

```{r ggplot plot simprof as dendro for more control}
hca <- res2$hclust %>% as.dendrogram %>% hang.dendrogram
clust <- cutree(hca,k=2)  
dendr <- dendro_data(hca, type = "rectangle")
clust.df <- data.frame(cluster = factor(clust))
clust.df$label <- rownames(clust.df)
dendr[["labels"]]   <- merge(dendr[["labels"]],clust.df, by="label")
rect <- aggregate(x~cluster,label(dendr),range)
rect <- data.frame(rect$cluster,rect$x)
ymax <- mean(hca$height[length(hca$height)-((k-2):(k-1))])


tiff("~/Documents/Chapter4/Oct2020_ICES/plots/Figure5_v3_2020-11-17.tiff", width=85,height=90, units = "mm", res = 300) 
ggplot() + 
  geom_segment(data=segment(dendr), aes(x=x, y=y, xend=xend, yend=yend),size = 0.3) + 
#  geom_text(data=label(dendr), aes(x, y,label=label, hjust=0, color=cluster), size=3, angle = 90) +
  geom_text(data=label(dendr), aes(x, y-13,label=c("Neutralized", "Untreated", "Neutralized", "Untreated", "Neutralized", "Untreated"), hjust=0, color=cluster), size=2.5, angle = 90) +
  scale_color_manual(values = c("maroon", "navy"))+
    annotate("text", label ="La Jolla\n Canyon", x = 1.5, y = -37, size = 2.5) +
  annotate("text", label = "CCE-2 Mooring", x = 4.5, y = -37, size = 2.5) + 
  annotate("text", label = "Day", x = 3.5, y = -25, size = 2.5) +
  annotate("text", label = "Night", x = 5.5, y = -25, size = 2.5) +
    geom_rect(aes(xmin=0.6, xmax=2.4, ymin=-45, ymax=50), 
            color=NA, fill=rgb(0,0,1, alpha = 0.2)) +
      geom_rect(aes(xmin=2.6, xmax=6.4, ymin=-45, ymax=50), 
            color=NA, fill=rgb(1,0,0, alpha = 0.2)) +
  theme_minimal(base_size = 6) +
    ylim(-50, 100) + 
  xlim(0.5,7) +
  scale_y_continuous(breaks=seq(-60,100,by=20), labels=c("","", "", "", "20", "40", "60", "80", "100")) +
#  coord_flex_cart(left=brackets_vertical())+
    ylab("Bray-Curtis Distance") + 
  xlab(NULL) +
  #geom_segment(aes(x = 0, y = 100, xend = 0, yend = 40)) +
  theme(legend.position = "none",
        axis.ticks.y.left = element_line(),
        axis.line.y.left = element_line(),
        plot.title = element_text(hjust = 0.5),
        axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),
        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        axis.text.y = element_text(color = "black", size = 5),  
        axis.title.y = element_text(color = "black", size = 6)) + 
  ggtitle("Bray-Curtis Distance, SIMPROF")+
  coord_capped_cart(left = "bottom") 
dev.off()

p + coord_flex_cart(left=brackets_vertical())
p + coord_capped_cart(gap = 1)
p + scale_y_continuous(expand = expand_scale(mult = c(0, 0)))



p
dev.off()

#ggdendrogram(res2dend, rotate = FALSE, size = 2) + geom_rect(data=rect, aes(xmin=X1-.3, xmax=X2+.3, ymin=0, ymax=ymax), color="red", fill=NA)+

```


```{r}
scratch <- rarefy_even_depth(COI_mergedbiomass)
bc_dist <- phyloseq::distance(scratch, method="bray", type = "samples")
sampledf <- data.frame(sample_data(scratch))
adonis(bc_dist ~ preservative*towlocation, data = sampledf)

ordination <- ordinate(scratch, method="MDS", "bray")
plot_ordination(scratch, ordination, color="preservative") + theme(aspect.ratio=1) + theme_bw()

adonis(bc_dist ~ sample_data(scratch)$preservative)

#pdf('plots/MDS_braycurtis.pdf', width=10,height=8) 
#plot_ordination(scratch, ordination, color="preservative", shape = "vial", size = 3) + theme(aspect.ratio=1) + theme_bw()
#dev.off()

```

Adonis R^2 tells us how much of the observed differences can be explained by each variable. towlocation (which here includes time and is unique for each tow) is the only significant variable out of the two, and explains 79% of the observed differences. preservative and the interaction between preservative and tow location are both non-significant, and explain only 2% and 2.5% of the observed differences, respectively. We can conclude that preservative has no effect. 



